---
title: "Gov 50 Final Project"
author: "Jason Rodrigues"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to how media coverage of the election changes the behavior of electorates. 




Have you ever wondered how national news coverage affects elections? I plan to examine how media coverage affects the electorate's behavior in this study. I hypothesize that national news media coverage will likely substantially influence the electorate's behavior. I affirm this because most citizens believe everything they get from news sources, and by believing the information, they are adopting the same political beliefs as the news source, ultimately shifting the election outcome. News sources can also push the political agenda by telling one side of a story, which electorates sympathize with and conform to the propaganda the sources are trying to push. However, this acutely insinuates that all news sources are bad, which is incredibly untrue. It can shine a light on candidates' ideals or actions, changing how electorates view the candidate they support. The explanatory variable is whether the town got Fox News; which is under the column of subrf 2000. and the gopvoteshare columns are those who voted for the Republican political party. The data supports my hypothesis that media coverage affects the electorate's behavior because it is seen through the data that a town that had Fox News subscribers has shifted towards the Republican party as elections went on. After all, the percentage of the town voting right-wing grew or increased. Subsequently, this affirms that the media affects the behavior of the electorate.




```{r}

library(tidyverse)
library(readr)
fox_news <- read_csv("foxnews.csv")

foxnews_plot <- fox_news |>
  
  ggplot(data = fox_news, mapping = aes(x = gopvoteshare2000, y = subrf2000,)) + geom_point(position = "dodge") + labs(x = "Percent of those that voted the GOP party in 2020", y = "Share of those who had FOX news subscriptions")
foxnews_plot
```
-- make an ayasis of the graph


```{r}
fox_news_plot_2 <- fox_news |>
  ggplot(data = fox_news, mapping = aes(x = gopvoteshare1996, y = subrf2000,)) + geom_point(position = "dodge") + labs(x = "Percent of those that voted the GOP party in 2020", y = "Share of those who had FOX news subscriptions")
fox_news_plot_2
```
From the scatter plot it is seen that with fox news the percentage that voted gop increased.

