---
title: "Gov 50 Final Project"
author: "Jason Rodrigues"
description: "Anaysis of what happens  to electorates behavior when Fox News is Introduced"
output:
  distill::distill_article:
    self_contained: false
---



## Introduction

  Have you ever wondered how national news coverage affects elections? This question instigated me to investigate the correlation between national news coverage and which political party electorates vote. To uncover the conclusion, I plan to examine how media coverage affects the electorate's behavior in this study. National news media coverage will likely substantially influence the electorate's behavior. I affirm this hypothesis because most citizens believe everything they get from news sources. By believing the information, they are adopting the same political beliefs as the news source, ultimately shifting the election's outcome. News sources can also push political propaganda by telling one side of a story, which electorates sympathize with and conform to the propaganda the sources are trying to push. However, this acutely insinuates that all news sources are bad, which is incredibly untrue. It can shine a light on candidates' ideals or actions, changing how electorates view the candidate they support. The type of news the electorates are watching is what primarily matters in this case due to the fact that news opinions and what is being shown varies among different news channels. To complete this analysis and hopefully uncover a clear answer, I decided to use electorate data from Fox News.
  
## The Data

   Fox News, if you did not already know, is a conservative media source that projects right-leaning opinion and news. Due to this, it makes it almost the perfect news station to investigate because since it clearly supports a side and the type of media it is going to project, I can use it for the project due to the fact I do not need to infer the opinion of Fox News. Also, the fact that this data set, in particular, only measures how their media affected the share of the town's population that voted for the GOP insinuates they will project conservative opinions and news--ultimately favoring the GOP. 

   Fox News was introduced in the late 1990s. Fox News had the same question, which prompted them to take this data. So, they decided to take data from  10,126 towns across 28 states in the United States to achieve a data set. 

   The explanatory variable is whether the town initially got Fox News or not, which is under the column of surf2000. Moreover, the gopvoteshare columns are those who voted for the Republican political party. The independent variable is the subr2000, which is the population share for each town that subscribes to Fox News. Then, the dependent variable is the percentage that voted for the GOP. A survey measures them. This experiment, in fact, is before and after. The before is the 1996 data when Fox News had just been introduced, and the after is the 2000 data when Fox News became more prominent and influential in media. The four years were enough difference in time for Fox News to push GOP propaganda. 


discuss the indpendant and dependant variable

discuss what type of design experiment is it

summarize the subr200 variable


## Data Visuals and Anaysis  


```{r Loading in the Data , echo=FALSE}

library(tidyverse)
library(readr)
fox_news <- read_csv("foxnews.csv")

```

Scatter plot #1

```{r, , echo=FALSE}
foxnews_plot <- fox_news |>
  ggplot(data = fox_news, mapping = aes(x = subrf2000, y = gopvoteshare1996,)) + geom_point(position = "dodge") + labs(x = "Share of those who had Fox News subscriptions", y = "Percent of those that voted the GOP party in 1996", title = "Correlation between Fox News subscription share and GOP voing share", subtitle = "Each point on the scatter plot repersents a town that Fox News collected data from in the 1996s") + geom_smooth() 
foxnews_plot

```

A scatter plot, in particular, is one of the first data recorded by Fox News. Fox News entered the media market around the late 1990s. The information for the data set uses the broad term of the late 1990s. However, with a quick Google search, I was able to find out that Fox News entered the media market around 1996. Due to this, this visualization is one of the first data sets. This graph may seem confusing due to the fact that all the points are clustered, and the best line of fit is wavy. However, it is more complex. The independent variable is the share of the population of the towns that had Fox News subscriptions. Due to cable company negotiations, Fox News was not able to enter conservative media first. Also, it explains why some towns are at zero because the town still needs Fox News available to them. The majority of the data rests between the Fox News share of 25% and 50%. However, it is quite difficult to establish a connection from this data because the best line of fit does not indicate an upward decline or a downward decline. Instead, it does both. However, if you shift your attention to the side where more than 50% of the town has Fox News, the best line of fit is able to tell us a story. The line curves up at a slow progression as the share of the population that obtained Fox News' subscriptions increases. Due to this, I am able to establish the fact that when more than 50 percent of the population share has Fox News, there is a positive causal relationship between the two variables; however, when less than 50% of the population has Fox News' subscriptions, the data is unclear from the scatter plot to establish a causal relationship.





Scatter plot #2

```{r echo=FALSE}
fox_news_plot_2 <- fox_news |>

  ggplot(data = fox_news, mapping = aes(x = subrf2000, y = gopvoteshare2000,)) + geom_point(position = "dodge") + labs(x = "Share of those who had Fox News subscriptions", y = "Percent of those that voted the GOP party in 2000", title = "Correlation between Fox News subscription share and GOP voing share", subtitle = "Each point on the scatter plot repersents a town that Fox News collected data from in the 2000") + geom_smooth() 
fox_news_plot_2
```
Looking at this scatter that has the GOP voting percentage from the year 2000 and the population share that had Fox News' subscriptions, there is a stark resemblance from the year 1996. This scatter plot represents the next election year, with Fox News replicating the way they took data from the last presidential election(in 1996 (which happens every four years)). The majority of the plots still fall between 25 percent and 50% of the share of the population that had Fox News. What is interesting here is that in the clump of points, when compared to the 1996 clump on the scatter plot above, it is seen that even though the best line of fit does not tell a clear story, Fox News had a conservative effect over time. This is because the data point clump in 2000 is slightly higher up on the y-axis compared to the data point clump in 1996. , considering no big events happened in the meantime, that the conservative jump was due to Fox News.

Furthermore, turning our attention back to the graph, when the share of the population reaches more than 50 percent, the curve makes a clear progressive shift upwards--except more exponentially when compared to the 1996 graph. Therefore, it is seen that when the population share is higher than 50%, there is a positive causal relationship. Although the best line of fit cannot tell us a story for the share that is lower than 50%, we can assume one thing. We can assume that throughout the four years, Fox News made many towns, on average, more conservative due to the data clump increasing on the y-axis. 

## Results (ATE)
```{r, echo=FALSE}
fox_news<- fox_news |> mutate(gopvotes2000mean = mean(gopvoteshare2000))


fox_news <- fox_news |> mutate(gopvotes1996mean = mean(gopvoteshare1996))

fox_news <- fox_news |> mutate(gopvotes1992mean = mean(gopvoteshare1992))



gopvotes2000 <- fox_news |> summarize(gopvotes2000mean = mean(gopvoteshare2000, gopvotes1996mean = mean(gopvoteshare1996)))



gopvotes1996 <- fox_news |> summarize(gopvotes1996mean = mean(gopvoteshare1996))



ATE <- fox_news |>
  group_by(gopvotes1996mean, gopvotes2000mean) |>
  summarize( ATE = 
gopvotes2000mean - gopvotes1996mean) |> slice(1:1)


knitr::kable(ATE, digits = 2)

 
```
This table is the ATE table. ATE stands for Average Treatment Affect. This method is very useful in this project because it is a simple table that will explain the differences between the 1996 GOP vote share percentage and the 2000 GOP share percentage from the presidential elections from that respective year. What I did was take the variables that reported the GOP share for each year, and I mutated and summarized a variable that averaged the two GOP percentages(mean). I did this so it can be modeled in a table. Now, looking at the table, it can be seen that I took the gopvotes2000mean and I minus the gopvotes1996mean and got the difference of 0.7. Since the ATE is positive, that means that the first variable, which was the 2000 mean, was larger than the 1996 mean. Thus, we are able to see that after four years, in 2000, on average, it got 0.07% more conservative with 0.07% more votes for the GOP.

## Main Regression 


```{r}
library(broom)
fox_news_main_regression <- 
lm(subrf2000 ~ gopvoteshare1996 + gopvoteshare2000, data = fox_news)

tidy(fox_news_main_regression) |>
knitr::kable(digits = 3)




```

This regression indicates that the coefficient, or the intercept, is  0.083. This means that with the implementation of Fox News, there was an increase of 0.083 in the percentage share of those in the town that voted GOP compared to towns that did not get Fox News. It is statistically significant at the test level of a = 0.005. We should interpret this casually. Since the p-value is 0, we would reject the null hypothesis that there is no association between Fox News subscriptions and GOP share increase.


## Conclusion

To summarize the data that I analyzed for this protection, I would like to revisit my hypothesis(look above in the introduction) and proclaim that I was correct. I said that media does have an effect on electorates behavior, and looking at the data, there sure is. Casual relationship. If you look at more than the scatter plots alone, then a relationship between a shift rightward and the implementation of Fox News is established. In my analysis of my graphs, I looked at the positive relationship between the increase in GOP share and the share of the population that subscribes to Fox News, which was not my hypothesis--rather, it was an interpretation of independent and dependent variables. I was testing if, over time, electorates' behavior changed due to the media, which, with the ATE table, established that voters got more conservative over time and voted for the GOP. There are a few limitations to my project. The first is that since there are so many points, they are not named, so I cannot see the direct change that the voting share changed over time on the 2000 scatter plot. The second limitation of my project is that I did not account for any omitted factors. This means that I did not adjust my data or my conclusion or graphs on any political events that happened in the four-year difference from when the data was taken. This data set, and even I ignore this factor. For instance, a stark political event could have happened that caused towns to shift rightward. If I had more time, I would have done outside research to see if there were any omitted variables that could have contributed to the shift. Moreover, if I had more time, I would have filtered my data set so that only 400 points showed instead of a thousand. I would want to do that because it will make it easier to read, and then I could label the points so that we can see the change in the voting share for the GOP over time. 




